\documentclass[11pt]{memoir}
\usepackage{longtable}
\usepackage{color}
\usepackage{tabu}
\usepackage{setspace}
\usepackage{pdflscape}
\usepackage{graphicx}
%\usepackage{subfigure}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{natbib}
\usepackage{fullpage}
%\bibliographystyle{plain}
\bibliographystyle{cbe}
\usepackage{algorithmic}
\usepackage[vlined,ruled]{algorithm2e}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{url}

\let \oriAlgorithm=\algorithm%
\renewcommand{\algorithm}[2]{%
	\oriAlgorithm%
	\SetAlgoLined%
	\DontPrintSemicolon%
	\SetKwComment{Comment}{$\triangleright\ $}{}%
	\TitleOfAlgo{ #1 }%
	\caption{ #2 }%
}%
\newcommand*{\Set}[2]{ #1 $\gets$ #2 \;}
\newcommand*{\SetC}[3]{ #1 $\gets$ #2 \Comment*[r]{#3}}
\newcommand*{\ForC}[3]{\For(\Comment*[f]{#2}){#1}{#3}}

\begin{document}
	%\firstpage{1}
	
	\title{Documentation for EUNCon Program}
	\author{Ward C. Wheeler\\
		Division of Invertebrate Zoology,\\ American Museum of Natural History,\\ 200 Central Park West, New York, NY, 10024, USA;\\wheeler@amnh.org}
	
	
	\maketitle
	\begin{center}
		Running Title: EUNCon
	\end{center}
	\newpage
	
	
	\section{Introduction}
	This is the first version of documentation for the program EUNCon.  This program is designed to produce a variety of 
	phylogenetic supergraphs from input graphs in a variety of formats and output the result in multiple formats as well.
	All source code, precompiled binaries, test data, and documentation are available from \url{https://githib.com/wardwheeler/EUNCon}.
	
	This first version is brief.
	
	\section{Input Graph Formats}
	Graphs may be input in the graphviz ``dot'' format \url{https://graphviz.org/}, Newick (as interpreted by Gary Olsen; \url{https://evolution.genetics.washington.edu/phylip/newick_doc.html}), Enhanced Newick \cite{Cardonaetal2008},
	and Forest Enhanced Newick (defined by \citealp{WheelerPhyloSuperGraphs}) formats.
	
	Quickly, Forest Enhanced Newick (FEN) is a format based on Enhanced Newick (Enewick) for forests of components, each
	of which is represented by an Enewick string.  The Enewick components are surrounded by `$<$' and '$>$'.
	As in $<$(A, (B,C)); (D,(E,F));$>$.  Groups may be shared among Enewick components.
	
	\section{Output Formats}
	Graph outputs can be in either Graphviz `dot' or FEN formats.  Dot files can be visualized in a variety of ways 
	using Graphviz (e.g. dot, neanto, twopi) into pdf, jpg and a large variety of other formats. FEN outputs of 
	single trees (ie forest with a single component) are rendered as enewick.  Newick files can be visualized in a 
	large number of programs (e.g. FigTree; \url{http://tree.bio.ed.ac.uk/software/figtree/})
	\bigskip
	\\Example dot command line: \\dot -Tpdf myDotFile.dot $>$ myDotFile.pdf\\
	
	For some reason on OSX the `pdf' option does not seem to work.  You can use `-Tps2' and that will generate a postscript file that preview can read and convert to pdf.
	
	\section{Command options}
	There are only a few program options that require specification.  There are defaults for all but input graphs.  
	Parameters are given with options in a range `a to b' (a-b) with any value in the interval, or alternates `a or b' (a|b). File options require a valid filename.
	For input graphs, wildcards are allowed (ie `*' and `?').  All commands are followed by a colon `:' before the option with no spaces.  Capitalization (for commands, but not filenames) is ignored.  Commands can be in any order (or entered from a file as stdin `$<$ filename).
	
	
	\begin{itemize}
		\item {Reconcile:eun$\mid$cun$\mid$majority$\mid$strict$\mid$Adams\\Default:eun\\This commands specifies the type of output graph.
		EUN is the Edge-Union-Network \cite{MiyagiandWheeler2019}, CUN the Cluster Union Network \citep{Baroni2005},
		majority (with fraction specified by `threshold') specifies that a values between 0 and 100 of either vertices or edges will be retained.  If all inputs are trees with the same leaf set this will be the Majority-Rule Consensus \citep{MargushandMcMorris1981}.
		Strict requires all vertices be present to be included in the final graph.  If all inputs are trees with the same leaf set this will be the Strict Consensus \citep{Schuhandpolhemus1980}. Adams denotes the Adams II consensus \citep{Adams1972}.}
		\item{Compare:Combinable$\mid$identity\\Default:combinable\\Species how group comparisons are to be made.  Either by identical match [(A, (B,C))$\neq$(A,B,C)],
			combinable sensu \cite{Nelson1979} [(A, (B,C)) consistent with (A,B,C)].  This option can be used to specify ``semi-strict'' consensus
			\citep{Bremer1990}.
		}
		\item{Threshold:(0-100)\\Default:0\\Threshold must be an integer between 0 and 100 and specifies the frequency of vertex or edge occurrence in input graphs to be included in the output graph.  Affects the behavior of `eun' and` majority.'
		}
		\item{Connect:True$\mid$False\\Default:False\\Specifies the output graph be connected (single component), potentially creating a root node and new edges labeled with ``0.0''.
		}
		\item{EdgeLabel:True$\mid$False\\Default:True\\Specifies the output graph have edges labeled with their frequency in input graphs.
		}
		\item{OutFormat:Dot$\mid$FENewick\\Default:Dot\\Specifies the output graph format as either Graphviz `dot' or FEN.
		}
		\item{OutFile:filename\\Default:euncon.out\\Specifies the output graph file name. No conventions are enforced.
		}
		\item{Any string that does not contain a colon, `:', is assumed to be an input graph file. 
		}
	\end{itemize}
	
	The program requires at least one input graph file and at least two input graphs (they could be in the same file).

	
	\section{Program Use}
	The program is invoked from the command-line as in:\\
	euncon reconcile:eun threshold:0 outfile:myOutput graphFile1.dot graphFile2.tre\\
	
	
	\subsection{Execution in Parallel}
	By default the program will execute using a single process core.  By specifying the options `+RTS -NX -RTS' where `X' is the number of processors offered to the program. These are specified after the program as in (for 4 parallel threads):\\
	\\
	euncon +RTS -N4 -RTS other options...  \\
	
	\section*{Acknowledgments}
	The author would like to thank NSF REU DBI-1358465, DARPA SIMPLEX N66001-15-C-4039, and Robert J. Kleberg Jr. and Helen C. Kleberg foundation grant ``Mechanistic Analyses of Pancreatic Cancer Evolution'' for financial support.  
	
	%\newpage
	%\bibliography{big-refs-3.bib}
	\bibliography{/users/ward/Dropbox/Work_stuff/manus/big-refs-3.bib}
	%\bibliography{/home/ward/Dropbox/Work_stuff/manus/big-refs-3.bib}
\end{document}